---
title: "jcomp2"
author: "Siddhant Mhapankar"
output: html_document
---

```{r}
library(ggplot2)
library(survival)
library(gbm)
library(caret)
library(pROC)
library(tree)
library(ISLR)
library(vip)
library(e1071)
library(tidyverse)
library(reshape2)
library(ggfortify)
library(rpart)
library(skimr)
library(corrplot)
library(kableExtra)
library(patchwork)
library(directlabels)
library(randomForest)
library(hrbrthemes)
library(viridis)
library(RColorBrewer)
heart<-read.csv("heart2.csv")
```

```{r}
head(heart)
```

```{r}
# Assign ID
heart$id <- seq.int(nrow(heart))

# Assign Character value to Numeric variables
heart$sexc <-ifelse(heart$sex==1, "Male", "Female")
heart$smoke <-ifelse(heart$smoking==1, "Yes", "No")
heart$hbp <- ifelse(heart$high_blood_pressure==1, "Yes","No")
heart$dia <-ifelse(heart$diabetes==1, "Yes", "No")
heart$anaemiac <- ifelse(heart$anaemia==1 ,"Yes", "No")
# Platelets : Hopkins Medicine
heart$platc <- ifelse(heart$platelets>150000 & heart$platelets <450000, "Platelets Normal", "Platelets Abnormal")
heart$plat <- ifelse(heart$platelets>150000 & heart$platelets <450000, 0,1)

# Serum Sodium: Mayo Clinic
heart$sodiumc <- ifelse(heart$serum_sodium >135 & heart$serum_sodium<145, "Serum Sodium Normal", "Serum Sodium Abnormal")
heart$sodiumn <- ifelse(heart$serum_sodium >135 & heart$serum_sodium<145, 0, 1)

#Creatine Phosphkinase : Mountsinai
heart$cpk <- ifelse(heart$creatinine_phosphokinase >10 & heart$creatinine_phosphokinase<120, "CPK Normal", "CPK Abnormal")
heart$cpkn <- ifelse(heart$creatinine_phosphokinase >10 & heart$creatinine_phosphokinase<120, 0, 1)

#ejection_fraction: Mayo
heart$efraction <-ifelse(heart$ejection_fraction<=75 & heart$ejection_fraction>=41, "Ejection Normal", "Ejection Abnormal")
heart$efractionn <-ifelse(heart$ejection_fraction<=75 & heart$ejection_fraction>=41, 0, 1)

#serum_creatinine :mayo
heart$screat<- ifelse((heart$serum_creatinine<1.35 & heart$serum_creatinine>0.74 & heart$sex==1 ) | (heart$serum_creatinine<1.04 & heart$serum_creatinine>0.59 & heart$sex==0) , "Creatinine Normal", "Creatinine Abnormal"   )
heart$screatn<- ifelse((heart$serum_creatinine<1.35 & heart$serum_creatinine>0.74 & heart$sex==1 ) | (heart$serum_creatinine<1.04 & heart$serum_creatinine>0.59 & heart$sex==0) , 0, 1 )

#age group: Pharma convention  
heart$agegp <- ifelse( heart$age<65, "Age <65", "Age >=65")
heart$agegpn <- ifelse( heart$age<65, 0, 1)

#event vs censor
heart$cnsr <- ifelse(heart$DEATH_EVENT==0, "Censor", "Event")
```

## Original Data table 
```{r}
h1<- subset(heart, select=c(age,anaemia,creatinine_phosphokinase, serum_creatinine,diabetes, ejection_fraction ,high_blood_pressure, platelets , serum_sodium, sex, smoking, DEATH_EVENT))
head(h1, 5)%>% DT::datatable()

```


## Training + Testing Data
```{r}
set.seed=8
train.test.split<-sample(2, nrow(h1), replace=TRUE, prob=c(0.8,0.2))
train=h1[train.test.split==1,]
test=h1[train.test.split==2,]

head(test, 5)%>% DT::datatable()
```

# Exploratory Data Analysis

## Binary Variable Distribution {.tabset .tabset-fade .tabset-pills}

```{r}
#1. age group
p1<-ggplot(heart, aes(x=agegp))+geom_bar(fill="lightblue")+ labs(x="Age Group")+ theme_minimal(base_size=10)

#2. Sex
p2<-ggplot(heart, aes(x=sexc))+geom_bar(fill="indianred3")+ labs(x="Sex")+ theme_minimal(base_size=10)

#3. Smoking
p3<-ggplot(heart, aes(x=smoke))+geom_bar(fill="seagreen2")+ labs(x="Smoking")+ theme_minimal(base_size=10)

#4. Diabetes
p4<-ggplot(heart, aes(x=dia))+geom_bar(fill="orange2")+
  labs(x="Diabetes Status")+ theme_minimal(base_size=10)

#5. cpk
p5<-ggplot(heart, aes(x=cpk))+geom_bar(fill="lightblue")+
labs(x="Creatinine Phosphokinase")+ theme_minimal(base_size=10)

#6. Platelets
p6<-ggplot(heart, aes(x=platc))+geom_bar(fill="indianred2")+
  labs(x="Platelets")+ theme_minimal(base_size=10)

#7. serum sodium
p7<-ggplot(heart, aes(x=sodiumc))+geom_bar(fill="seagreen2")+
  labs(x="Serum Sodium") + theme_minimal(base_size=10)

#8. Serum creatinine
p8<-ggplot(heart, aes(x=screat))+geom_bar(fill="orange2")+
  labs(x="Serum Creatinine") + theme_minimal(base_size=10)


#9. anaemia 
p9<-ggplot(heart, aes(x=anaemiac, fill=DEATH_EVENT))+geom_bar(fill="lightblue")+ labs(x="Anaemia")+ theme_minimal(base_size=10)

#10. ejection_fraction
p10<-ggplot(heart, aes(x=efraction))+geom_bar(fill="indianred2")+
  labs(x="Ejection Fraction")+ theme_minimal(base_size=10)

#11. High blood pressure
p11<-ggplot(heart, aes(x=hbp))+geom_bar(fill="seagreen2")+
  labs(x="High Blood Pressure Status")+ theme_minimal(base_size=10)

#12. Event
p12<-ggplot(heart, aes(x=cnsr))+geom_bar(fill="orangered3")+ labs(x="Event Status")+ theme_minimal(base_size=10)

```

### Demographic and Baseline Characters Distribution

```{r}
(p1+p2+p3 +p4)+
  plot_annotation(title="Demographic and Histology Distribution")
```


### Lab Test Result Distribution
```{r}
(p5+p6+p7+p8) + plot_annotation(title="Lab Test Distribution")
```


### Disease history Distribution
```{r}
(p9+p10+p11+p12) + plot_annotation(title="Disease History Distribution")
```


## Continuous Variables Disbribution  {.tabset .tabset-fade .tabset-pills}

### Age 
```{r}
#1. Age
c1<- ggplot(heart, aes(x=age))+ geom_histogram(binwidth=5, colour="white", fill="darkseagreen2", alpha=0.8)+
  geom_density(eval(bquote(aes(y=..count..*5))),colour="darkgreen", fill="darkgreen", alpha=0.3)+ scale_x_continuous(breaks=seq(40,100,10))+geom_vline(xintercept = 65, linetype="dashed")+ annotate("text", x=50, y=45, label="Age <65", size=2.5, color="dark green") + annotate("text", x=80, y=45, label="Age >= 65", size=2.5, color="dark red") +labs(title="Age Distribution") + theme_minimal(base_size = 8)
c1

```


### CPK

```{r}

#2. cpk
c2<- ggplot(heart, aes(x=creatinine_phosphokinase))+ geom_histogram(binwidth=100, colour="white", fill="mediumpurple2", alpha=0.8)+
  geom_density(eval(bquote(aes(y=..count..*150))),colour="mediumorchid1", fill="mediumorchid1", alpha=0.3)+ scale_x_continuous(breaks=seq(0,10000,1000))+geom_vline(xintercept = 120, linetype="dashed")+ annotate("text", x=0, y=100, label="CPK Normal", size=2.5, color="dark green") + annotate("text", x=1000, y=80, label="CPK Abnormal", size=2.5, color="dark red")+labs(title="Creatinine Phosphokinase Distribution") + theme_minimal(base_size = 8)
c2
```


### Ejection Fraction
```{r}
c3<- ggplot(heart, aes(x=ejection_fraction))+ geom_histogram(binwidth=5, colour="white", fill="lightpink1", alpha=0.8)+
  geom_density(eval(bquote(aes(y=..count..*5))),colour="mistyrose2", fill="mistyrose2", alpha=0.3)+ scale_x_continuous(breaks=seq(0,80,10))+geom_vline(xintercept = 40, linetype="dashed")+geom_vline(xintercept = 75, linetype="dashed")+ annotate("text", x=20, y=30, label="Abnormal", size=2.5, color="dark red") + annotate("text", x=50, y=30, label="Normal", color="dark green")+  annotate("text", x=80, y=30, label="Abnormal", size=2.5, color="dark red")+labs(title="Ejection Fraction Distribution") + theme_minimal(base_size = 8)
c3
```

### Platelets Count
```{r}
c4<- ggplot(heart, aes(x=platelets))+ geom_histogram(binwidth=20000, colour="white", fill="lightskyblue2", alpha=0.8)+
  geom_density(eval(bquote(aes(y=..count..*25000))),colour="lightsteelblue", fill="lightsteelblue", alpha=0.3)+
  geom_vline(xintercept = 150000, linetype="dashed")+geom_vline(xintercept = 450000, linetype="dashed")+ annotate("text", x=100000, y=30, label="Abnormal", size=2.5, color="dark red") + annotate("text", x=300000, y=30, label="Normal", color="dark green")+  annotate("text", x=500000, y=30, label="Abnormal", size=2.5, color="dark red")+labs(title="Platelets Count") + theme_minimal(base_size = 8)
c4
```



### Serum Sodium
```{r}
c5<- ggplot(heart, aes(x=serum_sodium))+ geom_histogram(binwidth=1, colour="white", fill="lightsalmon", alpha=0.8)+
  geom_density(eval(bquote(aes(y=..count..))),colour="lightcoral", fill="lightcoral", alpha=0.3)+
  geom_vline(xintercept = 135, linetype="dashed")+geom_vline(xintercept = 145, linetype="dashed")+ annotate("text", x=130, y=20, label="Abnormal", size=2.5, color="dark red") + annotate("text", x=142, y=20, label="Normal", color="dark green")+  annotate("text", x=148, y=20, label="Abnormal", size=2.5, color="dark red")+labs(title="Serum Sodium") + theme_minimal(base_size = 8)
c5
```


### Serum Creatinine
```{r}
c6<- ggplot(heart, aes(x=serum_creatinine))+ geom_histogram(binwidth=0.2, colour="white", fill="lightgoldenrod", alpha=0.8)+
  geom_density(eval(bquote(aes(y=..count..*0.2))),colour="moccasin", fill="moccasin", alpha=0.3)+
  geom_vline(xintercept = 0.74, linetype="dashed")+geom_vline(xintercept = 1.35, linetype="dashed")+ annotate("text", x=0.05, y=20, label="Abnormal", size=2.5, color="dark red") + annotate("text", x=1, y=20, label="Normal", color="dark green")+  annotate("text", x=2.5, y=20, label="Abnormal", size=2.5, color="dark red")+labs(title="Serum Creatinine") + theme_minimal(base_size = 8)
c6
```

## Correlations  {.tabset .tabset-fade .tabset-pills}

### Correlation Matrix

From the correlation matrix, we can see Death Event is highly correlated with serum creatinine, age, serum sodium, ejection fraction.

```{r}

r=cor(h1)
corrplot(r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 90)
```

### Heatmap

We will also show the heatmap as a side evidence.
```{r}
coul <- colorRampPalette(brewer.pal(8, "PiYG"))(25)
heatmap(r, scale="column", col = coul)
```

```{r}
draw_confusion_matrix <- function(cm) {

  total <- sum(cm$table)
  res <- as.numeric(cm$table)

  # Generate color gradients. Palettes come from RColorBrewer.
  greenPalette <- c("#F7FCF5","#E5F5E0","#C7E9C0","#A1D99B","#74C476","#41AB5D","#238B45","#006D2C","#00441B")
  redPalette <- c("#FFF5F0","#FEE0D2","#FCBBA1","#FC9272","#FB6A4A","#EF3B2C","#CB181D","#A50F15","#67000D")
  getColor <- function (greenOrRed = "green", amount = 0) {
    if (amount == 0)
      return("#FFFFFF")
    palette <- greenPalette
    if (greenOrRed == "red")
      palette <- redPalette
    colorRampPalette(palette)(100)[10 + ceiling(90 * amount / total)]
  }

  # set the basic layout
  layout(matrix(c(1,1,2)))
  par(mar=c(2,2,2,2))
  plot(c(100, 345), c(300, 450), type = "n", xlab="", ylab="", xaxt='n', yaxt='n')
  title('CONFUSION MATRIX', cex.main=2)

  # create the matrix 
  classes = colnames(cm$table)
  rect(150, 430, 240, 370, col=getColor("green", res[1]))
  text(195, 435, classes[1], cex=1.2)
  rect(250, 430, 340, 370, col=getColor("red", res[3]))
  text(295, 435, classes[2], cex=1.2)
  text(125, 370, 'Predicted', cex=1.3, srt=90, font=2)
  text(245, 450, 'Actual', cex=1.3, font=2)
  rect(150, 305, 240, 365, col=getColor("red", res[2]))
  rect(250, 305, 340, 365, col=getColor("green", res[4]))
  text(140, 400, classes[1], cex=1.2, srt=90)
  text(140, 335, classes[2], cex=1.2, srt=90)

  # add in the cm results
  text(195, 400, res[1], cex=1.6, font=2, col='white')
  text(195, 335, res[2], cex=1.6, font=2, col='white')
  text(295, 400, res[3], cex=1.6, font=2, col='white')
  text(295, 335, res[4], cex=1.6, font=2, col='white')

  # add in the specifics 
  plot(c(100, 0), c(100, 0), type = "n", xlab="", ylab="", main = "DETAILS", xaxt='n', yaxt='n')
  text(10, 85, names(cm$byClass[1]), cex=1.2, font=2)
  text(10, 70, round(as.numeric(cm$byClass[1]), 3), cex=1.2)
  text(30, 85, names(cm$byClass[2]), cex=1.2, font=2)
  text(30, 70, round(as.numeric(cm$byClass[2]), 3), cex=1.2)
  text(50, 85, names(cm$byClass[5]), cex=1.2, font=2)
  text(50, 70, round(as.numeric(cm$byClass[5]), 3), cex=1.2)
  text(70, 85, names(cm$byClass[6]), cex=1.2, font=2)
  text(70, 70, round(as.numeric(cm$byClass[6]), 3), cex=1.2)
  text(90, 85, names(cm$byClass[7]), cex=1.2, font=2)
  text(90, 70, round(as.numeric(cm$byClass[7]), 3), cex=1.2)

  # add in the accuracy information 
  text(30, 35, names(cm$overall[1]), cex=1.5, font=2)
  text(30, 20, round(as.numeric(cm$overall[1]), 3), cex=1.4)
  text(70, 35, names(cm$overall[2]), cex=1.5, font=2)
  text(70, 20, round(as.numeric(cm$overall[2]), 3), cex=1.4)
}

```

```{r}
library("gbm")
library("caret")
gbm.m<- gbm(train$DEATH_EVENT ~. , data=train, distribution = "bernoulli",
            cv.folds=10, shrinkage=0.01, n.minobsinnode = 10, n.trees=1000)
#gbm.m
gbm.imp=summary(gbm.m)
gbm.imp

gmb.t =predict(object=gbm.m, newdata=test, n.trees=1000, type="response")
presult<- as.factor(ifelse(gmb.t>0.5,1,0))
test$DEATH_EVENT1<-as.factor(test$DEATH_EVENT)
g<-confusionMatrix(presult,test$DEATH_EVENT1)
draw_confusion_matrix(g)

```

```{r}
paste("Ejection Fraction:  Normal 41%- 75%, Abnormal <41% or >75%")
a = heart$ejection_fraction
mean(a)
paste("The mean ejection fraction is in the abnormal level")
paste("Normal levels of serum creatinine 0.74 to 1.35 mg/dL")
b = heart$serum_creatinine
mean(b)
paste("The mean serum creatinine over the normal level")
paste("A normal platelet count ranges from 150,000 to 450,000 platelets per microliter of blood")
c = heart$platelets
mean(c)
paste("The mean platelet count is within the normal range")
paste("Two out of the three major factors lie in the abnormal level, hence the lifestyle of the citizens is poor.")


```

